---
  title: 'Exercises-1: Intro to R - NAs'
author: "Nathan Green, Imperial College London"
date: "07/09/2019"
output:
  pdf_document: default
html_document: default
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


First read in the data

```{r}
library(dataPakistan)
system.file("extdata", package = "dataPakistan") %>% dir()
```

# C:/Users/ngreen1/Documents/R/win-library/3.6/dataPakistan/extdata/Admin-datasheet-year2018.xlsx

ggplot(data = dat, mapping = aes(x = `Coverage at Hosehold (n)`, y = `Coverage at Hosehold (%)`)) + geom_point() + xlim(0, 1e+6)


my_plot <- ggplot(data = dat, mapping = aes(x = `Coverage at Hosehold (n)`, y = `Coverage at Hosehold (%)`))

my_plot + geom_point() + xlim(0, 1e+6)

###
install.packages("hexbin")
library(hexbin)

my_plot + geom_hex(bins = 50) + xlim(0, 1e+6)

###
my_plot + geom_point(alpha = 0.1) + xlim(0, 1e+6)

###
my_plot + geom_point(alpha = 0.1, colour = "blue") + xlim(0, 1e+6)

###
my_plot + geom_point(alpha = 0.1, aes(colour = Province)) + xlim(0, 1e+6)

###
ggplot(data = dat,
       mapping = aes(x = Province, y = `Coverage at Hosehold (n)`)) +
  geom_point(aes(colour = as.factor(Year))) + ylim(0, 1e+6)

###
ggplot(data = dat,
       mapping = aes(x = Province, y = `Coverage at Hosehold (n)`)) +
  geom_boxplot() + ylim(0, 1e+6)

###
ggplot(data = dat,
       mapping = aes(x = Province, y = `Coverage at Hosehold (n)`)) +
  geom_boxplot(alpha = 0) +
  geom_jitter(alpha = 0.3, colour = "tomato") +
  ylim(0, 1e+6)

###
ggplot(data = dat,
       mapping = aes(x = Province, y = `Coverage at Hosehold (n)`)) +
  geom_jitter(alpha = 0.3, colour = "tomato") +
  geom_boxplot(alpha = 0) +
  ylim(0, 1e+6)


# time series -------------------------------------------------------------

ggplot(data = dat, aes(x = Year, y = `Coverage at Hosehold (n)`)) + geom_line()

ggplot(data = dat[dat$District == "ATTOCK", ], aes(x = Year, y = `Coverage at Hosehold (n)`)) + geom_line()

ggplot(data = dat, aes(x = Year, y = `Coverage at Hosehold (n)`, colour = District)) + geom_line() + ylim(0, 1e+6)


# faceting ----------------------------------------------------------------

nms <- table(dat$District) %>% names %>% .[1:9]
ss_dat <- dat[dat$District %in% nms, ]

ggplot(data = ss_dat, mapping = aes(x = Year, y = `Coverage at Hosehold (n)`)) +
  geom_line() +
  facet_wrap(facets = vars(District))

ggplot(data = ss_dat, mapping = aes(x = Year, y = `Coverage at Hosehold (n)`)) +
  geom_line() +
  facet_wrap(facets = vars(District)) +
  theme_bw()


# arranging and exporting plots ------------------------------------------

install.packages("gridExtra")
library(gridExtra)

p1 <-
  ggplot(data = ss_dat, mapping = aes(x = Year, y = `Coverage at Hosehold (n)`)) +
  geom_line() +
  facet_wrap(facets = vars(District))

p2 <-
  ggplot(data = ss_dat, mapping = aes(x = Year, y = `Coverage at Hosehold (n)`)) +
  geom_line() +
  facet_wrap(facets = vars(District)) +
  theme_bw()

final_plot <- grid.arrange(p1, p2, ncol = 2)

ggsave("final_plot.png", final_plot, width = 10, dpi = 300)
