---
html_document: default
author: "Nathan Green, Imperial College London"
date: "07/09/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
title: 'Exercises: Joins and melts'
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


First load packages and read in the data

```r
library(dplyr)
library(ggplot2)
library(dataPakistan)

file_location <- system.file("extdata", package = "dataPakistan")

dat1 <- readxl::read_excel(path = paste0(file_location, "/MPQA _March SNID.xlsx"))

dat2 <- readxl::read_excel(path = paste0(file_location, "/LQAS %Pass _ District level_ Sep18-Mar19.xlsx"))
```

1. Both data sets have Province and District in them. Perform an inner join on these using `merge`.


```{r eval=FALSE}
xx = merge(x=dat1, y=dat2)
```

What happens and how can you fix it?

```{r eval=FALSE, echo = FALSE}
names(dat1)[names(dat1) == "Province name"] = "Province"
names(dat1)[names(dat1) == "District name"] = "District"
xx = merge(x=dat1, y=dat2)
```

After joining correctly what do you see?
Explain.

2. Perform a left join using `merge`.

```{r eval=FALSE, echo = FALSE}
xx = merge(x=dat12, y=dat22, all.x=T)
```

3. Perform a right join using `merge`
```{r eval=FALSE, echo = FALSE}
xx = merge(x=dat12, y=dat22, all.y=T)
```

4. Perform an outer join using `merge`.

```{r eval=FALSE, echo = FALSE}
xx = merge(x=dat1, y=dat2, all.x=TRUE, all.y=TRUE)
```

5. Now using `dplyr` we will do some filter joins.

Load the package

```{r}
suppressPackageStartupMessages(library(dplyr))
```

Perform an anti-join using.

```{r eval=FALSE, echo = FALSE}
xx = anti_join(x = dat1, y=dat2)
```

Which Province has the most missing entries?

6. Perform a semi-join.

```{r eval=FALSE, echo = FALSE}
xx = semi_join(x = dat1, y=dat2)
```

7. Melt `dat1` so that Province and District are the id columns and the others are the measure vaiables which form the variable and value columns.

```{r}
library(reshape2)
```

```{r eval = FALSE, echo=FALSE}
molten_dat <- melt(dat1, id.vars = c("Province", "District"))
```

8. Cast this long array so that it sums across District using `dcast()`.

```{r eval=FALSE, echo=FALSE}
dcast(molten_dat, Province ~ variable, fun.aggregate = sum, na.rm=TRUE)
```

