R provides many functions to examine features of vectors and other objects, for example

- `class()` - what kind of object is it (high-level)
- `typeof()` - what is the object’s data type (low-level)
- `length()` - how long is it? What about two dimensional objects
- `attributes()` - does it have any metadata

1. Type `x <- "dataset"` and `y <- 1:10`. What are the type and attributes of `x` and `y`?


```{r}
x <- "dataset"
y <- 1:10

typeof(x)

attributes(x)

y

typeof(y)

length(y)

z <- as.numeric(y)
z

typeof(z)
```

\noindent R has many data structures. These include

- atomic vector
- list
- matrix
- data frame
- factors
- Vectors

A vector is the most common and basic data structure in R and is pretty much the workhorse of R. Technically, vectors can be one of two types:

- atomic vectors
- lists

although the term “vector” most commonly refers to the atomic types not to lists.

### The Different Vector Modes
A vector is a collection of elements that are most commonly of mode character, logical, integer or numeric.

You can create an empty vector with `vector()`. (By default the mode is logical. You can be more explicit as shown in the examples below.) It is more common to use direct constructors such as `character()`, `numeric()`, etc.

2. Create a character vector of length 5.

```{r}
vector("character", length = 5)
#a vector of mode 'character' with 5 elements
```

3. Create character, numeric, logical vectors directly. What do you see?

```{r}
character(5) # the same thing, but using the constructor directly

numeric(5)   # a numeric vector with 5 elements

logical(5)   # a logical vector with 5 elements
```

\noindent You can also create vectors by directly specifying their content. R will then guess the appropriate mode of storage for the vector. For instance:

`x <- c(1, 2, 3)`
will create a vector x of mode numeric. These are the most common kind, and are treated as double precision real numbers. If you wanted to explicitly create integers, you need to add an L to each element (or coerce to the integer type using as.integer()).

4. Create a vector of 3 integers.

```{r}
x1 <- c(1L, 2L, 3L)
```

Using TRUE and FALSE will create a vector of mode logical:

`y <- c(TRUE, TRUE, FALSE, FALSE)`
While using quoted text will create a vector of mode character:

`z <- c("Sarah", "Tracy", "Jon")`

### Examining Vectors

The functions `str()` provide useful information about your vectors and R objects in general.

5. View the structure of `z` above.

### Adding Elements
The function `c()` (for combine) can also be used to add elements to a vector.

6. Attach `Annette` to the end of `z`. Attach `Greg` to the start.

```{r}
z <- c("Sarah", "Tracy", "Jon")
z <- c(z, "Annette")
z
z <- c("Greg", z)
z
```


### Vectors from a Sequence of Numbers

You can create vectors as a sequence of numbers using `:`.

7. Create a series from 1 to 10 using `:`. Create the same output using `seq()`. Change the step size to 0.1. 

```{r}
series <- 1:10
seq(10)
seq(from = 1, to = 10, by = 0.1)
```

8. Create the sequence above in reverse from 10 to 1.

```{r}
seq(from = 10, to = 1, by = -0.1)
```