## The console

Type the following with the correct operator

1. 3 plus 10
2. 6 times 12
3. What is 4 to the power 7?

```{r}
3 + 10
6 * 12
4^7
```

4. Type 3 + then
press `Enter`,
press `Enter`,
press `Enter`!

Whats happening?

press `Esc`


## Assigning values

1. Assign the value 100 to `x` using two different syntax

```{r}
x = 100
x <- 100
```

2. Assign the value 21 times `x` to `y` and print to console in one line of code.

```{r}
(y <- 21*x)
```

3. Now change the value of `x` to 200. What happens to the value of `y`?

```{r}
x <- 200
y
# y doesn't change i.e. its not a pointer
```


## Functions

1. Take the logarithm (natural base) of 10 by assigning 10 as an argument and calling the correct function. What value does it return?

```{r}
log(10)
```

2. What other log base functions are there?

```{r}
log10(10)
log2(10)
```

3. Take the reverse function to get back to 10.

```{r}
exp(log(10))
```

HINT: use the help documentation in the Help pane by calling `?` or `help()`


## Multiple argument

1. Use the `rep()` function to repeat the same numbers in a vector.
HINT: Look at the help document.

```{r}
rep(1)
```

2. Output to the console the number 1 five times.

```{r}
rep(1, 5)
```

3. Write the function with the argument for the number of times to repeat (5) first and 1 second.

```{r}
rep(times = 5, x = 1)
```

4. Use the `rep()` function to repeat the numbers 1,2 five times.

```{r}
rep(c(1,2), 5)
```

5. Change the arguments so that the total length of the output vector is length 5 with alternating numbers.

```{r}
rep(c(1,2), length = 5)
```

6. Change the arguments so that the total length of the output vector is length 5 but with all 1's together and all 2's together.

```{r}
rep(c(1,2), each = 3, length.out = 5)
```


## Variables

1. Which of these are numbers? Which are words? How can you tell?

```
1
"1"
"one"
```

```{r}
# number: 1
# words: "1", "one"
```


## Vectors

1. How many dimensions does a vector have?

```{r}
# 1
```

2. Consider a vector:

`x <- c(4,6,5,7,10,9,4,15)`

What is the value of:

`c(4,6,5,7,10,9,4,15) < 7`

a. `TRUE, FALSE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE`

b. `TRUE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE`

c. `FALSE, TRUE, TRUE, FALSE, FALSE, FALSE, TRUE, FALSE`

d. `TRUE, TRUE, TRUE, TRUE, TRUE, FALSE, TRUE, FALSE`


```{r}
# b
```

 
3. Consider two vectors:

`p <- c (3, 5, 6, 8)`
and

`q <- c (3, 3, 3)`

What is the value of:

`p+q`

a. `6, 8, 6, 8`

b. `6, 8, 0, 0`

c. `6, 8, NA, NA`

d. `3, 5, 6, 8 Warning message: In p+q : longer object length is not a multiple of shorter object length`

e. `6, 8, 9, 11`

```{r}
# e
```

4. If

`Age <- c(22, 25, 18, 20)`
`Name <- c("James", "Mathew", "Olivia", "Stella")`
`Gender <- c("M", "M", "F", "F")`

then what is the R code for getting the following output

```
##   Age   Name Gender
## 1  22  James      M
## 2  25 Mathew      M
```

a.

`DataFrame = data.frame(c(Age), c(Name), c(Gender))`

`subset(DataFrame, Gender == "M")`

b.

`DataFrame = data.frame(c(Age),c(Name),c(Gender))`

`subset(Gender=="M"), eval=FALSE` 

c.

`DataFrame = data.frame(Age,Name,Gender)`

`subset(DataFrame,Gender=="M")`

d.

`DataFrame = data.frame(c(Age,Name,Gender))`

`subset(DataFrame,Gender=="M")`
 

```{r}
# c
```


5. If `z <- 0:9` then what is the output from the following R-statements:

```
digits <- as.character(z)
as.integer(digits)
```

a. `Error in subset. object 'z' not found`

b. `0, 1, 2, 3, 4, 5, 6, 7, 8, 9`

c. `"NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA"`

d. `"0", "1", "2", "3", "4", "5", "6", "7", "8", "9"`

e. `0, 0, 0, 0, 0, 0, 0, 0, 0`

```{r}
# b
```


6. Consider the vector `x <- c(1,2,3,4)`.
What is the value of `k` for:

`(x+2)[(!is.na(x)) & x > 0] -> k`

a. `1, 2, 3, 4`

b. `1, 4, 9, 16`

c. `Error: object 'k' not found`

d. `3, 4, 5, 6`

e. `numeric(0)`

```{r}
# d
```

 
7. Consier the AirPassenger data set

`data(AirPassengers)`


Which statement will produce the following output?

`##  [1] 112 118 132 129 121 135 148 148 136 119 104 118`

a. `AirPassengers[time(AirPassengers) >= 1949 & time(AirPassengers) < 1950, 12]`

b. `AirPassengers[AirPassengers >= 1949 & AirPassengers < 1950]`

c. `AirPassengers[time(AirPassengers) >= 1949 & time(AirPassengers) < 1950]`

d. `AirPassengers[AirPassengers >= 1949 & AirPassengers < 1950, 12]`

e. `c[[1]]`

```{r}
# c
```


8. If `x <- c(2, 4, 6, 8)` and `y <- c(TRUE, TRUE, FALSE, TRUE)`

What is the value of `sum(x[y])`?

a. `20`

b. `8`

c. `14`

d. `NA`

```{r}
# c
```


9. Consider the vector `x <- c(34, 56, 55, 87, NA, 4, 77, NA, 21, NA, 39)`.

Which R-statement will count the number of NA values in x?

a. `count(is.na(X))`

b. `length(is.na(x))`

c. `sum(is.na(x))`

d. `count(!is.na(x))`

e. `sum(!is.na(x))`

```{r}
# c
```


10. How many dimensions does a matrix have?

```
1 2 3 4 5
2 3 4 5 6
3 4 5 6 7
4 5 6 7 8
```

```{r}
# 2
```

11. If the below matrix is named $M$ what is value $M_{34}$? How do you print it to screen?

```
1 2 3 4 5
2 3 4 5 6
3 4 5 6 7
4 5 6 7 8
```

```{r}
# 6
```

[https://www.r-exercises.com/2015/10/09/vector-exercises/]
