Read in the data

```{r}
library(dplyr)
library(dataPakistan)

data_name <- system.file(package = "dataPakistan", "extdata") %>% dir(full.names = TRUE) %>% .[2]
dat <- readxl::read_excel(data_name, sheet = 1) %>% as.data.frame()
```

First of all lets take a quick look at the data for a sanity check.

```{r}
dim(dat)
head(dat)
str(dat)
```

\noindent Its important to know what class each column has for when we do arithmetic and plotting witht them R will handle them differently.
We can check what types each column has using the `class()` function.

We can check each column in a simple loop (other ways are possible!)

```{r}
for (i in 1:ncol(dat)){
  print(typeof(dat[1,i]))
}
```

We see that the first 3 columns are characters and the rest numbers which seems sensible.

Now lets check for missing values. We can do this in the same way as above with a loop

```{r}
for (i in 1:ncol(dat)){
  print(anyNA(dat[,i]))
}
```

Looks good, no missing data.
Now we can check for some unusual values that may be typos are need further investigation.
Lets generate summarys for each column.

```{r}
for (i in 1:ncol(dat)){
  print(names(dat)[i])
  print(summary(dat[,i]))
}
```


The first column is months so lets check that there are only 12 of them using `table()`. This counts the frequencies.

```{r}
table(dat[ ,1])
```

Notice that the order is alphabetic and not starting from January. If we come to plot this data we may want to change this using `factors` and `levels`.
If we reorder then we can see if there are some missing months.

```{r}
x <- factor(dat[,1], levels = c("Jan","Feb","Mar","Apr","May","Jun","July","Aug","Sept","Oct","Nov","Dec"))
table(x)
```

This may need looking into further.