
Load in the `mtcars` data by typing `data("mtcars")`
This is described as Motor Trend Car Road Tests data.

Rather than refering to `mtcars` everytime we want to use data from it, we can `attach` the variable to the environment so that we can refer to the column names directly.
Do this by typing

```{r}
attach(mtcars)
```

Now we can simply type a columns name from those available i.e. `r names(mtcars)`.

1.
Plot Miles/(US) gallon versus Rear axle ratio by `plot(mpg,drat)`.
On which axis does `mpg` appear?
a. `mpg` appears on the x axis
b. `mpg` appears on the y axis

2.
Is a scatterplot recommended for continuous or dichotomous variables?
a. continuous
b. dichotomous

3
We want to include the origin on the plots. To do this we use the argument `xlim = c(0,<xmax>)` and `ylim = c(0,<ymax>)`.
Replot the scatterplot with the origin.

```{r}
plot(x=mpg, y=drat, xlim = c(0,35), ylim = c(0,5))
```

4 Let's change the type of points. We do this using the `pch` argument.

Redo the scatterplot with triangle points.
HINT: read the `pch` help document.

```{r}
plot(x=mpg, y=drat, xlim = c(0,35), ylim = c(0,5), pch = 2)
```

5. Give the plots a title using the `main` argument.

6. Give the plot a subtitle using `sub`.

7. Label the axes using `xlab` and `ylab`.

8. Make the size of the point half the size using `cex`

```{r}
plot(x=mpg, y=drat, xlim = c(0,35), ylim = c(0,5), cex = 0.5)
```

9. Change the colour of the points to dark blue.

```{r}
plot(x=mpg, y=drat, xlim = c(0,35), ylim = c(0,5), col = "dark blue")
```

10 If we set the `type` of plot as an argument to `'n'` then an empty plot is made. Create an empty scatterplot.
Then we can add points. Using the `text` function in the same way as `plot` (and `points`) plot a number for each point. HINT: with no `labels` argument in `text` sequential numbers are used.

```{r}
plot(x=mpg, y=drat, xlim = c(0,35), ylim = c(0,5), type = 'n')
text(x=mpg, y=drat)
```

11.
Produce a histogram with `hist(gear)`.
What do you see?
a. frequencies
b. probability density

12.
Change type of visualization of our scatterplot in Exercise 1 `plot(mpg,drat,type="")`.
If we want to see lines what we have to type into "":
a. `type="l"`
b. `type="p"`

13.
Now we want to see both point and lines in our plot.
What we have to type into `plot(mpg,drat,type="")`.
a. `type=c("p","l")`
b. `type="b"`

14.
Add another variable to our plot, for example `Weight`.
What command do we have to use:

a. `plot(mpg, drat); plot(mpg, wt)`
b. `plot(mpg, drat); points(mpg, wt)`

15.
Now we have added a new variable to our plot.
Suppose we want to use two different colours to separate the points.
Type `plot(mpg, drat, col=2)`.
What colour have we selected:

a. red
b. green

16.
Now we want to differentiate the two different variables in the scatterplot:

a. Let's change the colours of the second plot
b. Change use two different types of plot (e.g. points,lines)

17.
Now we want to highlight a variable in the final plot.
Type: `plot(mpg, drat, lwd=2) ; points(mpg, wt, lwd=1)`.
Which plot is highlighted:

a. plot1 (`mpg`, `drat`)
b. plot2 (`mpg`, `wt`)

18.
Finally choose four different continuous variables from mtcars set and produce:

a. Plot with lines and points for different variables with different colours (HINT: change y axis parameters by adding command `ylim=c(0,30)` to plot [e.g. `plot(a,b,type="p",ylim=c(0,30))`].

b. Choose one variable from each and highlighted it set red colour and a broad line.

[https://www.r-exercises.com/2015/10/09/vector-exercises/]
